---
title: "ponding boostrap test"
output: html_document
---

get the data ready for analysis
```{r}
library(readxl)
gas_dat <- read_excel("ponding_ghg_figshare.xlsx", 
    sheet = "dissolved_gases", na = "NA")
View(gas_dat)

##negative gas concentrations --> force everything below zero to zero
library(tidyr)
library(dplyr)
min(na.omit(gas_dat$pco2_uatm))   ## -146.87
gas_dat = gas_dat %>% mutate(pco2_uatm = if_else(pco2_uatm < 0, 0, pco2_uatm))
min(na.omit(gas_dat$pco2_uatm))
min(na.omit(gas_dat$pch4_uatm))  ##40.99 -- no need to change anything

#take the mean of the two center samples
centerdata = subset(gas_dat, samp_loc_in_pond == "center")

centerdata1 <- centerdata %>%
  group_by(date, pond, samp_round) %>%
  summarize(pco2_uatm = mean(pco2_uatm, na.rm = TRUE),
            pch4_uatm = mean(pch4_uatm, na.rm = TRUE))
head(centerdata1)

#add in a coumn to identify center as location in pond
centerdata1$samp_loc_in_pond = "center"

#shrink down the other dataset for just the data needed:
not_center = subset(gas_dat, samp_loc_in_pond != "center")

new_gas_dat = rbind(centerdata1, not_center)

#ok, so now need to squish things down by date
new_gas_means = new_gas_dat %>%
  group_by(date,pond, samp_round) %>%
  summarize(pco2_mean = mean(pco2_uatm, na.rm = TRUE),
            pch4_mean = mean(pch4_uatm, na.rm = TRUE))
new_gas_means
###looks good...
```

Make dataset for the temporal bootstraps:
```{r}
#start with the "new_gas_means"
new_gas_means

#make a "grand mean by pond" dataframe
library(tidyr)
library(dplyr)

big_means <- new_gas_means %>%
  group_by(pond) %>%
  summarize(bigco2mean = mean(pco2_mean, na.rm = TRUE),
            bigch4mean = mean(pch4_mean, na.rm = TRUE))

temporal_dat = merge(new_gas_means, big_means, by = c("pond"))

###need to remove Mud Pond -- only sampled twice...
temporal_dat = subset(temporal_dat, pond != "Mud Pond")
```

Ok, run bootstrap for pco2 temporal
```{r}
###be sure to run the lines below through the end of the "loop" all together (highlight before running)
sample_coef_intercept <- NULL
sample_coef_x1 <- NULL

library(dplyr)
new_df <- temporal_dat %>% group_by(pond) %>% slice_sample(n=1)

set.seed(2021)
for (i in 1:1000) {
  #Creating a resampled dataset from the sample data
  sample_d = temporal_dat %>% group_by(pond) %>% slice_sample(n=1)
  
  #Running the regression on these data
  model_bootstrap <- lm(pco2_mean ~ bigco2mean, data = sample_d)
  
  #Saving the coefficients
  sample_coef_intercept <-
    c(sample_coef_intercept, model_bootstrap$coefficients[1])
  
  sample_coef_x1 <-
    c(sample_coef_x1, model_bootstrap$coefficients[2])
}

summary(model_bootstrap)

#mean slope
mean_m = mean(sample_coef_x1)
#mean intercept
mean_int = mean(sample_coef_intercept)

mean_m
mean_int

#95% confidence intervals
a <-
  cbind(
    quantile(sample_coef_intercept, prob = 0.025),
    quantile(sample_coef_intercept, prob = 0.975))
b <-
  cbind(quantile(sample_coef_x1, prob = 0.025),
        quantile(sample_coef_x1, prob = 0.975))
a  #95% CI for the intercept
b  #95% CI for the slope

###now just find the max and min for the slope (or basically 100% CI)
c <-
  cbind(quantile(sample_coef_x1, prob = 0),
        quantile(sample_coef_x1, prob = 1))
c
```


Repeat for pCH4 over time
```{r}
###be sure to run the lines below through the end of the "loop" all together (highlight before running)
sample_coef_intercept <- NULL
sample_coef_x1 <- NULL

set.seed(2021)
for (i in 1:1000) {
  #Creating a resampled dataset from the sample data
  sample_d = temporal_dat %>% group_by(pond) %>% slice_sample(n=1)
  
  #Running the regression on these data
  model_bootstrap <- lm(pch4_mean ~ bigch4mean, data = sample_d)
  
  #Saving the coefficients
  sample_coef_intercept <-
    c(sample_coef_intercept, model_bootstrap$coefficients[1])
  
  sample_coef_x1 <-
    c(sample_coef_x1, model_bootstrap$coefficients[2])
}

summary(model_bootstrap)

#mean slope
mean_m = mean(sample_coef_x1)
#mean intercept
mean_int = mean(sample_coef_intercept)

mean_m
mean_int

#95% confidence intervals
a <-
  cbind(
    quantile(sample_coef_intercept, prob = 0.025),
    quantile(sample_coef_intercept, prob = 0.975))
b <-
  cbind(quantile(sample_coef_x1, prob = 0.025),
        quantile(sample_coef_x1, prob = 0.975))
a
b

###now just find the max and min for the slope (or basically 100% CI)
c <-
  cbind(quantile(sample_coef_x1, prob = 0),
        quantile(sample_coef_x1, prob = 1))
c
```



Make the spatial variance dataset
```{r}
#need to match up samples with mean on given date (so replicate new_gas_means three x and add center, and two non-center all in as one column...)
#center data already exists as "centerdata1"
#repeat for "out1 and out2"
out1dat = subset(gas_dat, samp_loc_in_pond == "out1")
out2dat = subset(gas_dat, samp_loc_in_pond == "out2")

out1dat$samp_loc_in_pond = "out1"

outdata1 <- out1dat %>%
  group_by(date, pond, samp_round, samp_loc_in_pond) %>%
  summarize(pco2_uatm = mean(pco2_uatm, na.rm = TRUE),
            pch4_uatm = mean(pch4_uatm, na.rm = TRUE))
head(outdata1)

out2dat$samp_loc_in_pond = "out2"

outdata2 <- out2dat %>%
  group_by(date, pond, samp_round, samp_loc_in_pond) %>%
  summarize(pco2_uatm = mean(pco2_uatm, na.rm = TRUE),
            pch4_uatm = mean(pch4_uatm, na.rm = TRUE))
head(outdata2)

center_space_dat = merge(centerdata1, new_gas_means, by = c ("pond", "date", "samp_round"))
out1_space_dat = merge(outdata1, new_gas_means, by = c ("pond", "date", "samp_round"))
out2_space_dat = merge(outdata2, new_gas_means, by = c ("pond", "date", "samp_round"))

spatial_dataset = rbind(center_space_dat, out1_space_dat, out2_space_dat)
#since no means on a given day at a given pond are the same, this setup should work...

##ok, now remove Mud Pond and Gibson Pond (only sampled from pond center)
spatial_dataset = subset(spatial_dataset, pond != "Gibson Pond")
spatial_dataset = subset(spatial_dataset, pond != "Mud Pond")
```


Try this approach for spatial CO2 first (based off of: https://towardsdatascience.com/bootstrap-regression-in-r-98bfe4ff5007)
```{r}
library(lme4)
library(lmerTest)
library(sjstats)
library(fitdistrplus)

###check distributions of mean pCO2 and all the individual pco2 samples
x = as.vector(na.omit(spatial_dataset$pco2_uatm + 1))
fit.lnorm.o2 = fitdist(x, "lnorm", method="mme")
fit.norm.o2 = fitdist(x, "norm", method="mme")
c(fit.norm.o2$aic, fit.lnorm.o2$aic)
#normal for individual samples

max(spatial_dataset$pco2_uatm)
min(spatial_dataset$pco2_uatm)

x = as.vector(na.omit(spatial_dataset$pco2_mean + 1))
fit.lnorm.o2 = fitdist(x, "lnorm", method="mme")
fit.norm.o2 = fitdist(x, "norm", method="mme")
c(fit.norm.o2$aic, fit.lnorm.o2$aic)
#also normal...

max(spatial_dataset$pco2_mean)
min(spatial_dataset$pco2_mean)


#be sure to run the lines below through "summary(model_bootstrap)" all together
sample_coef_intercept <- NULL
sample_coef_x1 <- NULL

set.seed(2021)
for (i in 1:1000) {
  #Creating a resampled dataset from the sample data
  sample_d = spatial_dataset %>% group_by(pond,samp_round) %>% slice_sample(n=1)#spatial_dataset[sample(1:nrow(spatial_dataset), nrow(spatial_dataset), replace = TRUE), ]
  
  #Running the regression on these data
  #model_bootstrap <- lmer(pco2_uatm ~ pco2_mean + (1|pond), data = sample_d)
  model_bootstrap = lm(pco2_uatm ~ pco2_mean, data = sample_d)
  
  #Saving the coefficients
  #sample_coef_intercept <- c(sample_coef_intercept, model_bootstrap@beta[1])
  #sample_coef_x1 <- c(sample_coef_x1, model_bootstrap@beta[2])
  sample_coef_intercept <- c(sample_coef_intercept, model_bootstrap$coefficients[1])
  sample_coef_x1 <- c(sample_coef_x1, model_bootstrap$coefficients[2])
}

summary(model_bootstrap)
r2(model_bootstrap)

#mean slope
mean_m = mean(sample_coef_x1)
#mean intercept
mean_int = mean(sample_coef_intercept)
mean_m
mean_int

#confidence intervals
a <-
  cbind(
    quantile(sample_coef_intercept, prob = 0.025),
    quantile(sample_coef_intercept, prob = 0.975))
b <-
  cbind(quantile(sample_coef_x1, prob = 0.025),
        quantile(sample_coef_x1, prob = 0.975))
a
b ###these two values are the low and high slopes within the 95% confidence interval

c <-
  cbind(quantile(sample_coef_x1, prob = 0),
        quantile(sample_coef_x1, prob = 1))
c
```



And for spatial ch4
```{r}
sample_coef_intercept <- NULL
sample_coef_x1 <- NULL

set.seed(2021)
for (i in 1:1000) {
  #Creating a resampled dataset from the sample data
  sample_d = spatial_dataset %>% group_by(pond,samp_round) %>% slice_sample(n=1)#spatial_dataset[sample(1:nrow(spatial_dataset), nrow(spatial_dataset), replace = TRUE), ]
  
  #Running the regression on these data
  #model_bootstrap <- lmer(pco2_uatm ~ pco2_mean + (1|pond), data = sample_d)
  model_bootstrap = lm(pch4_uatm ~ pch4_mean, data = sample_d)
  
  #Saving the coefficients
  #sample_coef_intercept <- c(sample_coef_intercept, model_bootstrap@beta[1])
  #sample_coef_x1 <- c(sample_coef_x1, model_bootstrap@beta[2])
  sample_coef_intercept <- c(sample_coef_intercept, model_bootstrap$coefficients[1])
  sample_coef_x1 <- c(sample_coef_x1, model_bootstrap$coefficients[2])
}

summary(model_bootstrap)
r2(model_bootstrap)

#mean slope
mean_m = mean(sample_coef_x1)
#mean intercept
mean_int = mean(sample_coef_intercept)
mean_m
mean_int

#confidence intervals
a <-
  cbind(
    quantile(sample_coef_intercept, prob = 0.025),
    quantile(sample_coef_intercept, prob = 0.975))
b <-
  cbind(quantile(sample_coef_x1, prob = 0.025),
        quantile(sample_coef_x1, prob = 0.975))
a
b ###these two values are the low and high slopes within the 95% confidence interval

c <-
  cbind(quantile(sample_coef_x1, prob = 0),
        quantile(sample_coef_x1, prob = 1))
c
```




Repeat for pCH4 over space and time (what happens if you grab a single sample all summer)
Make the dataset first
```{r}
#use big mean from temporal dataset
temporal_dat$bigch4mean
temporal_dat$bigco2mean

library(dplyr)

space_time_means = spatial_dataset %>%
  group_by(pond) %>%
  summarize(bigco2mean = mean(pco2_mean, na.rm = TRUE),
            bigch4mean = mean(pch4_mean, na.rm = TRUE))

space_time = merge(spatial_dataset, space_time_means, by = c("pond"))
```

Run it for CH4
```{r}
###be sure to run the lines below through the end of the "loop" all together (highlight before running, or run the whole chunk)
sample_coef_intercept <- NULL
sample_coef_x1 <- NULL

####need to change to the "Big Mean" and not just each individual date mean

set.seed(2021)
for (i in 1:1000) {
  #Creating a resampled dataset from the sample data
  sample_d = space_time %>% group_by(pond) %>% slice_sample(n=1)
  
  #Running the regression on these data
  model_bootstrap <- lm(pch4_uatm ~ bigch4mean, data = sample_d)
  
  #Saving the coefficients
  sample_coef_intercept <-
    c(sample_coef_intercept, model_bootstrap$coefficients[1])
  
  sample_coef_x1 <-
    c(sample_coef_x1, model_bootstrap$coefficients[2])
}

summary(model_bootstrap)

#mean slope
mean_m = mean(sample_coef_x1)
#mean intercept
mean_int = mean(sample_coef_intercept)
mean_m
mean_int

#95% confidence intervals
a <-
  cbind(
    quantile(sample_coef_intercept, prob = 0.025),
    quantile(sample_coef_intercept, prob = 0.975))
b <-
  cbind(quantile(sample_coef_x1, prob = 0.025),
        quantile(sample_coef_x1, prob = 0.975))
a
b

###now just find the max and min for the slope (or basically 100% CI)
c <-
  cbind(quantile(sample_coef_x1, prob = 0),
        quantile(sample_coef_x1, prob = 1))
c
```

and for pCO2
```{r}
###be sure to run the lines below through the end of the "loop" all together (highlight before running)
sample_coef_intercept <- NULL
sample_coef_x1 <- NULL

set.seed(2021)
for (i in 1:1000) {
  #Creating a resampled dataset from the sample data
   sample_d = space_time %>% group_by(pond) %>% slice_sample(n=1)
  
  #Running the regression on these data
  model_bootstrap <- lm(pco2_uatm ~ bigco2mean, data = sample_d)
  
  #Saving the coefficients
  sample_coef_intercept <-
    c(sample_coef_intercept, model_bootstrap$coefficients[1])
  
  sample_coef_x1 <-
    c(sample_coef_x1, model_bootstrap$coefficients[2])
}

summary(model_bootstrap)

#mean slope
mean_m = mean(sample_coef_x1)
#mean intercept
mean_int = mean(sample_coef_intercept)
mean_m
mean_int

#95% confidence intervals
a <-
  cbind(
    quantile(sample_coef_intercept, prob = 0.025),
    quantile(sample_coef_intercept, prob = 0.975))
b <-
  cbind(quantile(sample_coef_x1, prob = 0.025),
        quantile(sample_coef_x1, prob = 0.975))
a
b

###now just find the max and min for the slope (or basically 100% CI)
c <-
  cbind(quantile(sample_coef_x1, prob = 0),
        quantile(sample_coef_x1, prob = 1))
c
```

